---
title: "Pre- and post-fire anaylisis of soils" 
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(knitr)
library(rmdformats)

## Global options
options(max.print="100")
opts_chunk$set(echo=FALSE,
	             cache=FALSE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE,
               autodep = TRUE)
```

```{r pkgs}
library("tidyverse")
library("here")
library("janitor")
library("ggpubr")
library(fitdistrplus)
library(lme4)
library(lmerTest)
library(emmeans)
library(multcomp)
library(performance)
library(afex)
```

# Prepare data 
```{r prepare-data, echo=TRUE}
raw_soil <- readxl::read_excel(
  here::here("data/Resultados_Suelos_2018_2021_v2.xlsx"),
  sheet = "SEGUIMIENTO_SUELOS_sin_ouliers"
) %>% 
  janitor::clean_names() %>% 
  mutate(
    treatment_name =
      case_when(
        str_detect(geo_parcela_nombre, "NP_") ~ "Autumn Burning / No Browsing",
        str_detect(geo_parcela_nombre, "PR_") ~ "Spring Burning / Browsing",
        str_detect(geo_parcela_nombre, "P_") ~ "Autumn Burning / Browsing"
      ),
    zona =
      case_when(
        str_detect(geo_parcela_nombre, "NP_") ~ "QOt_NP",
        str_detect(geo_parcela_nombre, "PR_") ~ "QPr_P",
        str_detect(geo_parcela_nombre, "P_") ~ "QOt_P"
      ),
    fecha = lubridate::ymd(fecha),
    pre_post_quema = case_when(
      pre_post_quema == "Prequema" ~ "0 preQuema",
      pre_post_quema == "Postquema" ~ "1 postQuema"
    )
  ) 

```

- Select data pre- and intermediately post-fire (first post-fire sampling: "2018-12-20" and "2019-05-09" for autumn and spring fires respectively)

```{r, echo=TRUE}
soil <- raw_soil %>% 
  filter(fecha %in% lubridate::ymd(c("2018-12-11", "2018-12-20", "2019-04-24", "2019-05-09"))) %>% 
  mutate(zona = as.factor(zona), 
         pre_post_quema = as.factor(pre_post_quema)
  )
```

- Structure of the data

```{r}
xtabs(~pre_post_quema+zona, data = soil)
```

# Modelize 

- For each response variable, the approach modelling is 

$Y \sim zona (P|NP|PR) + Fecha(pre|post) + zona \times Fecha$ 

- using the "(1|zona:geo_parcela_nombre)" as nested random effects

```{r}
# mira esto 
# https://doi.org/10.1186/s42408-018-0022-8 
```
- Then explore error distribution of the variable response and model diagnostics 
- Select the appropiate error distribution and use LMM or GLMM 
- Explore Post-hoc  
- Plot interactions 


# Humedad 
### Model 
```{r}
# Normal 
yvar <- "humedad"
df_model <- soil %>% 
  dplyr::select(one_of(c(yvar,"pre_post_quema", "zona", "geo_parcela_nombre")))
form <- reformulate("pre_post_quema * zona + (1|zona:geo_parcela_nombre)", 
                    response = yvar)

m <- lme4::lmer(form, data = df_model)
anova(as_lmerModLmerTest(m))

tabla_anova <- as.data.frame(anova(as_lmerModLmerTest(m)))
tabla_anova$variable <- yvar
tabla_anova$factor <- rownames(tabla_anova)
tabla_anova

tabla_anova %>% dplyr::select(fvalue = `F value`, p=`Pr(>F)`, variable, factor) %>%  
  assign(paste0("anova_", yvar), ., inherits = TRUE)

```

### *Post-hoc*
```{r}
emmeans(m, list(pairwise ~ pre_post_quema), adjust = "tukey")
emmeans(m, list(pairwise ~zona), adjust = "tukey")
emmeans(m, list(pairwise ~ pre_post_quema | zona), adjust = "tukey")
```






```{r}
cols <- c("#619CFF","#F8766D","#00BA38")

          # "QOt_NP" = "#619CFF",
          # "QPr_P" = "#00BA38",
          # "QOt_P" = "#F8766D")

plot_interaction <- function(df, yvar){
  pos <- position_dodge(0.15)
  df %>%
    dplyr::select(UQ(yvar), 
                  pre_post_quema, 
                  zona) %>% 
    group_by(zona,pre_post_quema) %>% 
    summarise(mean = mean(.data[[yvar]], na.rm=FALSE),
              se = plotrix::std.error(.data[[yvar]], na.rm = FALSE)) %>% 
    ggplot(aes(x=pre_post_quema, y = mean, 
               group=zona, color = zona, fill=zona, shape=zona)) + 
    geom_point(position = pos, size = 3) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se), 
                  position = pos, width = .1) +
    geom_line(position = pos) +
    theme_pubr() + 
    ylab(yvar) + xlab("") +
    scale_color_manual(values = cols) + 
    scale_fill_manual(values=cols)
}
```


# CIC 
### Model 
```{r}
# Normal 
yvar <- "cic"
df_model <- soil %>% 
  dplyr::select(one_of(c(yvar,"pre_post_quema", "zona", "geo_parcela_nombre")))
form <- reformulate("pre_post_quema * zona + (1|zona:geo_parcela_nombre)", 
                    response = yvar)

m <- lme4::lmer(form, data = df_model)
anova(as_lmerModLmerTest(m))

tabla_anova <- as.data.frame(anova(as_lmerModLmerTest(m)))
tabla_anova$variable <- yvar
tabla_anova$factor <- rownames(tabla_anova)
tabla_anova

tabla_anova %>% dplyr::select(fvalue = `F value`, p=`Pr(>F)`, variable, factor) %>%  
  assign(paste0("anova_", yvar), ., inherits = TRUE)

```

### *Post-hoc*
```{r}
emmeans(m, list(pairwise ~ pre_post_quema), adjust = "tukey")
emmeans(m, list(pairwise ~zona), adjust = "tukey")
emmeans(m, list(pairwise ~ pre_post_quema | zona), adjust = "tukey")
```

















# C 
### Model 
```{r}
# Normal 
yvar <- "c_percent"
df_model <- soil %>% 
  dplyr::select(one_of(c(yvar,"pre_post_quema", "zona", "geo_parcela_nombre")))
form <- reformulate("pre_post_quema * zona + (1|zona:geo_parcela_nombre)", 
                    response = yvar)

m <- lme4::lmer(form, data = df_model)
anova(as_lmerModLmerTest(m))

tabla_anova <- as.data.frame(anova(as_lmerModLmerTest(m)))
tabla_anova$variable <- yvar
tabla_anova$factor <- rownames(tabla_anova)
tabla_anova

tabla_anova %>% dplyr::select(fvalue = `F value`, p=`Pr(>F)`, variable, factor) %>%  
  assign(paste0("anova_", yvar), ., inherits = TRUE)

```

### *Post-hoc*
```{r}
emmeans(m, list(pairwise ~ pre_post_quema), adjust = "tukey")
emmeans(m, list(pairwise ~zona), adjust = "tukey")
emmeans(m, list(pairwise ~ pre_post_quema | zona), adjust = "tukey")
```



















# Fe
### Model 
```{r}
# Gamma
yvar <- "fe_percent"
df_model <- soil %>% 
  dplyr::select(one_of(c(yvar,"pre_post_quema", "zona", "geo_parcela_nombre")))
form <- reformulate("pre_post_quema * zona + (1|zona:geo_parcela_nombre)", 
                    response = yvar)

m <- lme4::glmer(form, data = df_model, family= Gamma)
afex::mixed(form, data=df_model)
tabla_anova <- afex::mixed(form, data=df_model)
tabla_anova$anova_table
tabla_anova <- as.data.frame(tabla_anova$anova_table)
tabla_anova$variable <- yvar
tabla_anova$factor <- rownames(tabla_anova)

tabla_anova %>% dplyr::select(fvalue = `F`, p=`Pr(>F)`, variable, factor) %>%  
  assign(paste0("anova_", yvar), ., inherits = TRUE)

```

### *Post-hoc*
```{r}
emmeans(m, list(pairwise ~ pre_post_quema), adjust = "tukey")
emmeans(m, list(pairwise ~zona), adjust = "tukey")
emmeans(m, list(pairwise ~ pre_post_quema | zona), adjust = "tukey")
```



















# Fe
### Model 
```{r}
# Gamma
yvar <- "mo"
df_model <- soil %>% 
  dplyr::select(one_of(c(yvar,"pre_post_quema", "zona", "geo_parcela_nombre")))
form <- reformulate("pre_post_quema * zona + (1|zona:geo_parcela_nombre)", 
                    response = yvar)

m <- lme4::glmer(form, data = df_model, family= Gamma)
afex::mixed(form, data=df_model)
tabla_anova <- afex::mixed(form, data=df_model)
tabla_anova$anova_table
tabla_anova <- as.data.frame(tabla_anova$anova_table)
tabla_anova$variable <- yvar
tabla_anova$factor <- rownames(tabla_anova)

tabla_anova %>% dplyr::select(fvalue = `F`, p=`Pr(>F)`, variable, factor) %>%  
  assign(paste0("anova_", yvar), ., inherits = TRUE)

```

### *Post-hoc*
```{r}
emmeans(m, list(pairwise ~ pre_post_quema), adjust = "tukey")
emmeans(m, list(pairwise ~zona), adjust = "tukey")
emmeans(m, list(pairwise ~ pre_post_quema | zona), adjust = "tukey")
```



















# K
### Model 
```{r}
# Gamma
yvar <- "k_percent"
df_model <- soil %>% 
  dplyr::select(one_of(c(yvar,"pre_post_quema", "zona", "geo_parcela_nombre")))
form <- reformulate("pre_post_quema * zona + (1|zona:geo_parcela_nombre)", 
                    response = yvar)

m <- lme4::glmer(form, data = df_model, family= Gamma)
afex::mixed(form, data=df_model)
tabla_anova <- afex::mixed(form, data=df_model)
tabla_anova$anova_table
tabla_anova <- as.data.frame(tabla_anova$anova_table)
tabla_anova$variable <- yvar
tabla_anova$factor <- rownames(tabla_anova)

tabla_anova %>% dplyr::select(fvalue = `F`, p=`Pr(>F)`, variable, factor) %>%  
  assign(paste0("anova_", yvar), ., inherits = TRUE)

```

### *Post-hoc*
```{r}
emmeans(m, list(pairwise ~ pre_post_quema), adjust = "tukey")
emmeans(m, list(pairwise ~zona), adjust = "tukey")
emmeans(m, list(pairwise ~ pre_post_quema | zona), adjust = "tukey")
```



















# Mg
### Model 
```{r}
# Gamma
yvar <- "mg_percent"
df_model <- soil %>% 
  dplyr::select(one_of(c(yvar,"pre_post_quema", "zona", "geo_parcela_nombre")))
form <- reformulate("pre_post_quema * zona + (1|zona:geo_parcela_nombre)", 
                    response = yvar)

m <- lme4::glmer(form, data = df_model, family= Gamma)
afex::mixed(form, data=df_model)
tabla_anova <- afex::mixed(form, data=df_model)
tabla_anova$anova_table
tabla_anova <- as.data.frame(tabla_anova$anova_table)
tabla_anova$variable <- yvar
tabla_anova$factor <- rownames(tabla_anova)

tabla_anova %>% dplyr::select(fvalue = `F`, p=`Pr(>F)`, variable, factor) %>%  
  assign(paste0("anova_", yvar), ., inherits = TRUE)

```

### *Post-hoc*
```{r}
emmeans(m, list(pairwise ~ pre_post_quema), adjust = "tukey")
emmeans(m, list(pairwise ~zona), adjust = "tukey")
emmeans(m, list(pairwise ~ pre_post_quema | zona), adjust = "tukey")
```

# C/N
### Model 
```{r}
# Gamma
yvar <- "c_n"
df_model <- soil %>% 
  dplyr::select(one_of(c(yvar,"pre_post_quema", "zona", "geo_parcela_nombre")))
form <- reformulate("pre_post_quema * zona + (1|zona:geo_parcela_nombre)", 
                    response = yvar)

m <- lme4::glmer(form, data = df_model, family= Gamma)
afex::mixed(form, data=df_model)
tabla_anova <- afex::mixed(form, data=df_model)
tabla_anova$anova_table
tabla_anova <- as.data.frame(tabla_anova$anova_table)
tabla_anova$variable <- yvar
tabla_anova$factor <- rownames(tabla_anova)

tabla_anova %>% dplyr::select(fvalue = `F`, p=`Pr(>F)`, variable, factor) %>%  
  assign(paste0("anova_", yvar), ., inherits = TRUE)

```

### *Post-hoc*
```{r}
emmeans(m, list(pairwise ~ pre_post_quema), adjust = "tukey")
emmeans(m, list(pairwise ~zona), adjust = "tukey")
emmeans(m, list(pairwise ~ pre_post_quema | zona), adjust = "tukey")
```

# P
### Model 
```{r}
# Negative Binomial 
yvar <- "p"
df_model <- soil %>% 
  dplyr::select(one_of(c(yvar,"pre_post_quema", "zona", "geo_parcela_nombre")))
form <- reformulate("pre_post_quema * zona + (1|zona:geo_parcela_nombre)", 
                    response = yvar)

m <- lme4::glmer.nb(form, data = df_model, family= Gamma)
afex::mixed(form, data=df_model)
tabla_anova <- afex::mixed(form, data=df_model)
tabla_anova$anova_table
tabla_anova <- as.data.frame(tabla_anova$anova_table)
tabla_anova$variable <- yvar
tabla_anova$factor <- rownames(tabla_anova)

tabla_anova %>% dplyr::select(fvalue = `F`, p=`Pr(>F)`, variable, factor) %>%  
  assign(paste0("anova_", yvar), ., inherits = TRUE)

```

### *Post-hoc*
```{r}
emmeans(m, list(pairwise ~ pre_post_quema), adjust = "tukey")
emmeans(m, list(pairwise ~zona), adjust = "tukey")
emmeans(m, list(pairwise ~ pre_post_quema | zona), adjust = "tukey")
```

# N
### Model 
```{r}
# Beta
yvar <- "n_percent"
df_model <- soil %>% 
  dplyr::select(one_of(c(yvar,"pre_post_quema", "zona", "geo_parcela_nombre")))
form <- reformulate("pre_post_quema * zona + (1|zona:geo_parcela_nombre)", 
                    response = yvar)

library(glmmTMB)
m <- glmmTMB::glmmTMB(formula = form,
                         family = beta_family (link = "logit"),
                         data = df_model)

tabla_anova <- car::Anova(m)
tabla_anova$variable <- yvar
tabla_anova$factor <- rownames(tabla_anova)
assign(paste0("anova_", yvar), tabla_anova)

  tabla_anova %>% dplyr::select(fvalue = `Chisq`, p=`Pr(>Chisq)`, variable, factor) %>%  
  assign(paste0("anova_", yvar), ., inherits = TRUE)
```

### *Post-hoc*
```{r}
emmeans(m, list(pairwise ~ pre_post_quema), adjust = "tukey")
emmeans(m, list(pairwise ~zona), adjust = "tukey")
emmeans(m, list(pairwise ~ pre_post_quema | zona), adjust = "tukey")
```








# Na
### Model 
```{r}
# Beta
yvar <- "na_percent"
df_model <- soil %>% 
  dplyr::select(one_of(c(yvar,"pre_post_quema", "zona", "geo_parcela_nombre")))
form <- reformulate("pre_post_quema * zona + (1|zona:geo_parcela_nombre)", 
                    response = yvar)

library(glmmTMB)
m <- glmmTMB::glmmTMB(formula = form,
                         family = beta_family (link = "logit"),
                         data = df_model)

tabla_anova <- car::Anova(m)
tabla_anova$variable <- yvar
tabla_anova$factor <- rownames(tabla_anova)
assign(paste0("anova_", yvar), tabla_anova)

  tabla_anova %>% dplyr::select(fvalue = `Chisq`, p=`Pr(>Chisq)`, variable, factor) %>%  
  assign(paste0("anova_", yvar), ., inherits = TRUE)
```

### *Post-hoc*
```{r}
emmeans(m, list(pairwise ~ pre_post_quema), adjust = "tukey")
emmeans(m, list(pairwise ~zona), adjust = "tukey")
emmeans(m, list(pairwise ~ pre_post_quema | zona), adjust = "tukey")
```








# pH agua 
### Model 
```{r}
# Gamma
yvar <- "p_h_agua_eez"
df_model <- soil %>% 
  dplyr::select(one_of(c(yvar,"pre_post_quema", "zona", "geo_parcela_nombre")))
form <- reformulate("pre_post_quema * zona + (1|zona:geo_parcela_nombre)", 
                    response = yvar)

m <- lme4::glmer(form, data = df_model, family= Gamma)
afex::mixed(form, data=df_model)
tabla_anova <- afex::mixed(form, data=df_model)
tabla_anova$anova_table
tabla_anova <- as.data.frame(tabla_anova$anova_table)
tabla_anova$variable <- yvar
tabla_anova$factor <- rownames(tabla_anova)

tabla_anova %>% dplyr::select(fvalue = `F`, p=`Pr(>F)`, variable, factor) %>%  
  assign(paste0("anova_", yvar), ., inherits = TRUE)

```

### *Post-hoc*
```{r}
emmeans(m, list(pairwise ~ pre_post_quema), adjust = "tukey")
emmeans(m, list(pairwise ~zona), adjust = "tukey")
emmeans(m, list(pairwise ~ pre_post_quema | zona), adjust = "tukey")
```




# pH KCl 
### Model 
```{r}
# Gamma
yvar <- "p_h_k_cl"
df_model <- soil %>% 
  dplyr::select(one_of(c(yvar,"pre_post_quema", "zona", "geo_parcela_nombre")))
form <- reformulate("pre_post_quema * zona + (1|zona:geo_parcela_nombre)", 
                    response = yvar)

m <- lme4::glmer(form, data = df_model, family= Gamma)
afex::mixed(form, data=df_model)
tabla_anova <- afex::mixed(form, data=df_model)
tabla_anova$anova_table
tabla_anova <- as.data.frame(tabla_anova$anova_table)
tabla_anova$variable <- yvar
tabla_anova$factor <- rownames(tabla_anova)

tabla_anova %>% dplyr::select(fvalue = `F`, p=`Pr(>F)`, variable, factor) %>%  
  assign(paste0("anova_", yvar), ., inherits = TRUE)

```

### *Post-hoc*
```{r}
emmeans(m, list(pairwise ~ pre_post_quema), adjust = "tukey")
emmeans(m, list(pairwise ~zona), adjust = "tukey")
emmeans(m, list(pairwise ~ pre_post_quema | zona), adjust = "tukey")
```

# NH4 
- prepara datos 

```{r}
soilN <- raw_soil %>% 
  dplyr::select(zona, fecha, pre_post_quema, geo_parcela_nombre, 
                n_nh4, n_no3) %>% 
  drop_na() %>% 
  mutate(zona = as.factor(zona), 
         pre_post_quema = as.factor(pre_post_quema)
  )
  
xtabs(~pre_post_quema+zona, data = soilN)


soilN %>% 
  group_by(zona, pre_post_quema) %>% 
  summarise(N.nh4 = length(n_nh4),
            N.no3 = length(n_no3))
```
### Model 
```{r}
# Gamma
yvar <- "n_nh4"
df_model <- soilN %>% 
  dplyr::select(one_of(c(yvar,"pre_post_quema", "zona", "geo_parcela_nombre")))
form <- reformulate("pre_post_quema * zona + (1|zona:geo_parcela_nombre)", 
                    response = yvar)

m <- lme4::glmer(form, data = df_model, family= Gamma)
afex::mixed(form, data=df_model)
tabla_anova <- afex::mixed(form, data=df_model)
tabla_anova$anova_table
tabla_anova <- as.data.frame(tabla_anova$anova_table)
tabla_anova$variable <- yvar
tabla_anova$factor <- rownames(tabla_anova)

tabla_anova %>% dplyr::select(fvalue = `F`, p=`Pr(>F)`, variable, factor) %>%  
  assign(paste0("anova_", yvar), ., inherits = TRUE)

```

### *Post-hoc*
```{r}
emmeans(m, list(pairwise ~ pre_post_quema), adjust = "tukey")
emmeans(m, list(pairwise ~zona), adjust = "tukey")
emmeans(m, list(pairwise ~ pre_post_quema | zona), adjust = "tukey")
```









# NO3
### Model 
```{r}
# Gamma
yvar <- "n_no3"
df_model <- soilN %>% 
  dplyr::select(one_of(c(yvar,"pre_post_quema", "zona", "geo_parcela_nombre")))
form <- reformulate("pre_post_quema * zona + (1|zona:geo_parcela_nombre)", 
                    response = yvar)

m <- lme4::glmer(form, data = df_model, family= Gamma)
afex::mixed(form, data=df_model)
tabla_anova <- afex::mixed(form, data=df_model)
tabla_anova$anova_table
tabla_anova <- as.data.frame(tabla_anova$anova_table)
tabla_anova$variable <- yvar
tabla_anova$factor <- rownames(tabla_anova)

tabla_anova %>% dplyr::select(fvalue = `F`, p=`Pr(>F)`, variable, factor) %>%  
  assign(paste0("anova_", yvar), ., inherits = TRUE)

```

### *Post-hoc*
```{r}
emmeans(m, list(pairwise ~ pre_post_quema), adjust = "tukey")
emmeans(m, list(pairwise ~zona), adjust = "tukey")
emmeans(m, list(pairwise ~ pre_post_quema | zona), adjust = "tukey")
```


# General Overview
## Mean + SE table 
```{r}
library(plotrix)
library(gtsummary)
soil %>% 
  dplyr::select(zona, pre_post_quema, one_of("humedad","n_nh4","n_no3","fe_percent","k_percent","mg_percent","na_percent","n_percent","c_percent","c_n","cic","p","mo","p_h_k_cl","p_h_agua_eez")) %>% 
  tbl_strata(
    strata = pre_post_quema, 
    .tbl_fun = 
      ~.x %>% 
      tbl_summary(
        by = zona,
        missing = "no",
        type = c(p) ~ "continuous",
       # label = list(ab ~ "Juvenile abundance"),
        statistic = all_continuous() ~ "{mean} ({std.error})",
        digits = everything() ~ 2
      ))
```
## Figures 
```{r}
pos <- position_dodge(0.15)

medias <- soil %>% dplyr::select(
  pre_post_quema, zona, humedad:p_h_agua_eez)%>% 
  group_by(zona, pre_post_quema) %>% 
  summarise_at(vars(humedad:p_h_agua_eez), ~mean(., na.rm=TRUE)) %>% 
  pivot_longer(humedad:p_h_agua_eez, names_to = "var", values_to = "mean")

errores <- soil %>% dplyr::select(
  pre_post_quema, zona, humedad:p_h_agua_eez) %>% 
  group_by(zona, pre_post_quema) %>% 
  summarise_at(vars(humedad:p_h_agua_eez), ~plotrix::std.error(., na.rm=TRUE)) %>% 
  pivot_longer(humedad:p_h_agua_eez, names_to = "var", values_to = "se")

medias <- medias %>% dplyr::inner_join(errores) %>% 
  mutate(varname = case_when(
    var == "n_percent" ~ "N",
    var == "p" ~ "P",
    var == "k_percent" ~ "K",
    var == "mo" ~ "M.O.",
    var == "c_percent" ~ "C",
    var == "c_n" ~ "C/N",
    var == "humedad" ~ "Humedad",
    var == "n_nh4" ~ "NH4",
    var == "fe_percent" ~ "Fe",
    var == "mg_percent" ~ "Mg",
    var == "na_percent" ~ "Na",
    var == "cic" ~ "C.I.C.",
    var == "n_no3" ~ "NO3",
    var == "p_h_k_cl"  ~ "pH (KCl)",
    var == "p_h_agua_eez" ~ "pH (Agua)",
    var == "p_h_agua" ~ "pH (Agua)_old"
  ))

plotsuelos <- function(df, ncol, nrow){
    ggplot(df, aes(x=pre_post_quema, y = mean, 
               color = zona, shape = zona, group=zona)) + 
    geom_point(position = pos, size = 2) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se), 
                  position = pos, width = .01) +
    geom_line(position = pos) +
    theme_bw() +
    theme(strip.background = element_rect(fill="white")) + 
    ylab("") + xlab("") +
  facet_wrap(~varname, scales = "free_y", ncol=ncol)
}
```


```{r}
medias %>% 
  filter(var %in% c("n_percent", "p", "k_percent",
                    "mo", "c_percent", "c_n")) %>% 
  plotsuelos(ncol=2) 
```

```{r}
medias %>% 
  filter(var %in% c("cic", "mg_percent", 
                    "na_percent", "humedad")) %>% 
  plotsuelos(ncol=2)
```

```{r}
medias %>% 
  filter(var %in% c("n_nh4","n_no3",
                    "p_h_k_cl", "p_h_agua_eez")) %>% 
  plotsuelos(ncol=2) 
```

## Anovas table
```{r}

anovas <- bind_rows(anova_c_n, 
                    anova_cic,
                    anova_k_percent,
                    anova_mg_percent,
                    anova_mo,
                    anova_n_nh4,
                    anova_n_no3, 
                    anova_p, 
                    anova_p_h_agua_eez, 
                    anova_p_h_k_cl, 
                    anova_humedad, 
                    anova_n_percent, 
                    anova_c_percent, 
                    anova_na_percent) %>% 
  relocate(variable, factor)
rownames(anovas) <- NULL
```

```{r}
library(kableExtra)

anovas %>% 
  mutate(across(where(is.numeric), round, 3)) %>% 
  pivot_wider(names_from = factor, values_from = c(fvalue, p)) %>% 
  relocate(variable,
           fvalue_zona, p_zona, 
           fvalue_pre_post_quema, p_pre_post_quema, 
           `fvalue_pre_post_quema:zona`, `p_pre_post_quema:zona`
           ) %>% 
  kbl(col.names = 
        c("Variables", rep(c("F", "p"),3))) %>% 
  add_header_above(c("", "zona" = 2, "fecha" = 2, "zona X fecha" = 2)) %>% 
  kable_styling() 

```

```{r}
anovas %>% 
  mutate(across(where(is.numeric), round, 3)) %>% 
  pivot_wider(names_from = factor, values_from = c(fvalue, p)) %>% 
  relocate(variable,
           fvalue_zona, p_zona, 
           fvalue_pre_post_quema, p_pre_post_quema, 
           `fvalue_pre_post_quema:zona`, `p_pre_post_quema:zona`
           ) %>% 
  write_csv(here::here("output/anovas_pre_post.csv"))
```


