---
title: "Pre- and post-fire anaylisis of soils by Fire date" 
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(knitr)
library(rmdformats)

## Global options
options(max.print="100")
opts_chunk$set(echo=FALSE,
	             cache=FALSE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE,
               autodep = TRUE)
```

```{r pkgs}
library("tidyverse")
library("here")
library("janitor")
library("ggpubr")
library(fitdistrplus)
library(lme4)
library(lmerTest)
library(emmeans)
library(multcomp)
library(performance)
library(afex)
library(kableExtra)

```

# Prepare data 
```{r prepare-data, echo=TRUE}
raw_soil <- readxl::read_excel(
  here::here("data/Resultados_Suelos_2018_2021_v2.xlsx"),
  sheet = "SEGUIMIENTO_SUELOS_sin_ouliers"
) %>% 
  janitor::clean_names() %>% 
  mutate(
    treatment_name =
      case_when(
        str_detect(geo_parcela_nombre, "NP_") ~ "Autumn Burning / No Browsing",
        str_detect(geo_parcela_nombre, "PR_") ~ "Spring Burning / Browsing",
        str_detect(geo_parcela_nombre, "P_") ~ "Autumn Burning / Browsing"
      ),
    zona =
      case_when(
        str_detect(geo_parcela_nombre, "NP_") ~ "QOt_NP",
        str_detect(geo_parcela_nombre, "PR_") ~ "QPr_P",
        str_detect(geo_parcela_nombre, "P_") ~ "QOt_P"
      ),
    estacion =
      case_when(
        str_detect(geo_parcela_nombre, "NP_") ~ "Ot",
        str_detect(geo_parcela_nombre, "PR_") ~ "Pr",
        str_detect(geo_parcela_nombre, "P_") ~ "Ot"
      ),
    date = lubridate::ymd(fecha),
    fecha = case_when(
      pre_post_quema == "Prequema" ~ "0 preQuema",
      pre_post_quema == "Postquema" ~ "1 postQuema"
    )
  ) 

```

- Select data pre- and intermediately post-fire (first post-fire sampling: "2018-12-20" and "2019-05-09" for autumn and spring fires respectively)

```{r, echo=TRUE}
soil <- raw_soil %>% 
  filter(date %in% lubridate::ymd(c("2018-12-11", "2018-12-20", "2019-04-24", "2019-05-09"))) %>% 
  mutate(zona = as.factor(zona), 
         fecha = as.factor(fecha)
  )
```

- Structure of the data

```{r}
xtabs(~fecha+estacion, data = soil)
```

# Modelize 

- For each response variable, the approach modelling is 

$Y \sim estacion (Ot|Pr) + fecha(pre|post) + estacion \times fecha$ 

- using the "(1|estacion:geo_parcela_nombre)" as nested random effects

```{r}
# mira esto 
# https://doi.org/10.1186/s42408-018-0022-8 
```
- Then explore error distribution of the variable response and model diagnostics 
- Select the appropiate error distribution and use LMM or GLMM 
- Explore Post-hoc  
- Plot interactions 

```{r}
# custom posthocs

tabla_postHoc <- function(modelo){
  posthoc <- emmeans(modelo, ~ fecha * estacion)
  contrast(posthoc, "consec", simple = "each", combine = TRUE,
           adjust = "tukey") %>% 
  kbl(digits = 4) %>% 
    kable_paper("hover", full_width = F)
}


```


# Humedad 
### Model 
```{r}
# Normal 
yvar <- "humedad"
df_model <- soil %>% 
  dplyr::select(one_of(c(yvar,"fecha", "estacion", "geo_parcela_nombre")))
form <- reformulate("fecha * estacion + (1|estacion:geo_parcela_nombre)", 
                    response = yvar)

m <- lme4::lmer(form, data = df_model)
anova(as_lmerModLmerTest(m))

tabla_anova <- as.data.frame(anova(as_lmerModLmerTest(m)))
tabla_anova$variable <- yvar
tabla_anova$factor <- rownames(tabla_anova)
tabla_anova

tabla_anova %>% dplyr::select(fvalue = `F value`, p=`Pr(>F)`, variable, factor) %>%  
  assign(paste0("anova_", yvar), ., inherits = TRUE)

```

### *Post-hoc*
```{r}
tabla_postHoc(m)
```


# CIC 
### Model 
```{r}
# Normal 
yvar <- "cic"
df_model <- soil %>% 
  dplyr::select(one_of(c(yvar,"fecha", "estacion", "geo_parcela_nombre")))
form <- reformulate("fecha * estacion + (1|estacion:geo_parcela_nombre)", 
                    response = yvar)

m <- lme4::lmer(form, data = df_model)
anova(as_lmerModLmerTest(m))

tabla_anova <- as.data.frame(anova(as_lmerModLmerTest(m)))
tabla_anova$variable <- yvar
tabla_anova$factor <- rownames(tabla_anova)
tabla_anova

tabla_anova %>% dplyr::select(fvalue = `F value`, p=`Pr(>F)`, variable, factor) %>%  
  assign(paste0("anova_", yvar), ., inherits = TRUE)

```

### *Post-hoc*
```{r}
tabla_postHoc(m)
```


# C 
### Model 
```{r}
# Normal 
yvar <- "c_percent"
df_model <- soil %>% 
  dplyr::select(one_of(c(yvar,"fecha", "estacion", "geo_parcela_nombre")))
form <- reformulate("fecha * estacion + (1|estacion:geo_parcela_nombre)", 
                    response = yvar)

m <- lme4::lmer(form, data = df_model)
anova(as_lmerModLmerTest(m))

tabla_anova <- as.data.frame(anova(as_lmerModLmerTest(m)))
tabla_anova$variable <- yvar
tabla_anova$factor <- rownames(tabla_anova)
tabla_anova

tabla_anova %>% dplyr::select(fvalue = `F value`, p=`Pr(>F)`, variable, factor) %>%  
  assign(paste0("anova_", yvar), ., inherits = TRUE)

```

### *Post-hoc*
```{r}
tabla_postHoc(m)
```



# Fe
### Model 
```{r}
# Gamma
yvar <- "fe_percent"
df_model <- soil %>% 
  dplyr::select(one_of(c(yvar,"fecha", "estacion", "geo_parcela_nombre")))
form <- reformulate("fecha * estacion + (1|estacion:geo_parcela_nombre)", 
                    response = yvar)

m <- lme4::glmer(form, data = df_model, family= Gamma)
afex::mixed(form, data=df_model)
tabla_anova <- afex::mixed(form, data=df_model)
tabla_anova$anova_table
tabla_anova <- as.data.frame(tabla_anova$anova_table)
tabla_anova$variable <- yvar
tabla_anova$factor <- rownames(tabla_anova)

tabla_anova %>% dplyr::select(fvalue = `F`, p=`Pr(>F)`, variable, factor) %>%  
  assign(paste0("anova_", yvar), ., inherits = TRUE)

```

### *Post-hoc*
```{r}
tabla_postHoc(m)
```


# MO
### Model 
```{r}
# Gamma
yvar <- "mo"
df_model <- soil %>% 
  dplyr::select(one_of(c(yvar,"fecha", "estacion", "geo_parcela_nombre")))
form <- reformulate("fecha * estacion + (1|estacion:geo_parcela_nombre)", 
                    response = yvar)

m <- lme4::glmer(form, data = df_model, family= Gamma)
afex::mixed(form, data=df_model)
tabla_anova <- afex::mixed(form, data=df_model)
tabla_anova$anova_table
tabla_anova <- as.data.frame(tabla_anova$anova_table)
tabla_anova$variable <- yvar
tabla_anova$factor <- rownames(tabla_anova)

tabla_anova %>% dplyr::select(fvalue = `F`, p=`Pr(>F)`, variable, factor) %>%  
  assign(paste0("anova_", yvar), ., inherits = TRUE)

```

### *Post-hoc*
```{r}
tabla_postHoc(m)
```


# K
### Model 
```{r}
# Gamma
yvar <- "k_percent"
df_model <- soil %>% 
  dplyr::select(one_of(c(yvar,"fecha", "estacion", "geo_parcela_nombre")))
form <- reformulate("fecha * estacion + (1|estacion:geo_parcela_nombre)", 
                    response = yvar)

m <- lme4::glmer(form, data = df_model, family= Gamma)
afex::mixed(form, data=df_model)
tabla_anova <- afex::mixed(form, data=df_model)
tabla_anova$anova_table
tabla_anova <- as.data.frame(tabla_anova$anova_table)
tabla_anova$variable <- yvar
tabla_anova$factor <- rownames(tabla_anova)

tabla_anova %>% dplyr::select(fvalue = `F`, p=`Pr(>F)`, variable, factor) %>%  
  assign(paste0("anova_", yvar), ., inherits = TRUE)

```

### *Post-hoc*
```{r}
tabla_postHoc(m)
```



# Mg
### Model 
```{r}
# Gamma
yvar <- "mg_percent"
df_model <- soil %>% 
  dplyr::select(one_of(c(yvar,"fecha", "estacion", "geo_parcela_nombre")))
form <- reformulate("fecha * estacion + (1|estacion:geo_parcela_nombre)", 
                    response = yvar)

m <- lme4::glmer(form, data = df_model, family= Gamma)
afex::mixed(form, data=df_model)
tabla_anova <- afex::mixed(form, data=df_model)
tabla_anova$anova_table
tabla_anova <- as.data.frame(tabla_anova$anova_table)
tabla_anova$variable <- yvar
tabla_anova$factor <- rownames(tabla_anova)

tabla_anova %>% dplyr::select(fvalue = `F`, p=`Pr(>F)`, variable, factor) %>%  
  assign(paste0("anova_", yvar), ., inherits = TRUE)

```

### *Post-hoc*
```{r}
tabla_postHoc(m)
```

# C/N
### Model 
```{r}
# Gamma
yvar <- "c_n"
df_model <- soil %>% 
  dplyr::select(one_of(c(yvar,"fecha", "estacion", "geo_parcela_nombre")))
form <- reformulate("fecha * estacion + (1|estacion:geo_parcela_nombre)", 
                    response = yvar)

m <- lme4::glmer(form, data = df_model, family= Gamma)
afex::mixed(form, data=df_model)
tabla_anova <- afex::mixed(form, data=df_model)
tabla_anova$anova_table
tabla_anova <- as.data.frame(tabla_anova$anova_table)
tabla_anova$variable <- yvar
tabla_anova$factor <- rownames(tabla_anova)

tabla_anova %>% dplyr::select(fvalue = `F`, p=`Pr(>F)`, variable, factor) %>%  
  assign(paste0("anova_", yvar), ., inherits = TRUE)

```

### *Post-hoc*
```{r}
tabla_postHoc(m)
```


# P
### Model 
```{r}
# Negative Binomial 
yvar <- "p"
df_model <- soil %>% 
  dplyr::select(one_of(c(yvar,"fecha", "estacion", "geo_parcela_nombre")))
form <- reformulate("fecha * estacion + (1|estacion:geo_parcela_nombre)", 
                    response = yvar)

m <- lme4::glmer.nb(form, data = df_model, family= Gamma)
afex::mixed(form, data=df_model)
tabla_anova <- afex::mixed(form, data=df_model)
tabla_anova$anova_table
tabla_anova <- as.data.frame(tabla_anova$anova_table)
tabla_anova$variable <- yvar
tabla_anova$factor <- rownames(tabla_anova)

tabla_anova %>% dplyr::select(fvalue = `F`, p=`Pr(>F)`, variable, factor) %>%  
  assign(paste0("anova_", yvar), ., inherits = TRUE)

```

### *Post-hoc*
```{r}
tabla_postHoc(m)
```

# N
### Model 
```{r}
# Beta
yvar <- "n_percent"
df_model <- soil %>% 
  dplyr::select(one_of(c(yvar,"fecha", "estacion", "geo_parcela_nombre")))
form <- reformulate("fecha * estacion + (1|estacion:geo_parcela_nombre)", 
                    response = yvar)

library(glmmTMB)
m <- glmmTMB::glmmTMB(formula = form,
                         family = beta_family (link = "logit"),
                         data = df_model)

tabla_anova <- car::Anova(m)
tabla_anova$variable <- yvar
tabla_anova$factor <- rownames(tabla_anova)
assign(paste0("anova_", yvar), tabla_anova)

  tabla_anova %>% dplyr::select(fvalue = `Chisq`, p=`Pr(>Chisq)`, variable, factor) %>%  
  assign(paste0("anova_", yvar), ., inherits = TRUE)

car::Anova(m)
```

### *Post-hoc*
```{r}
tabla_postHoc(m)
```

# Na
### Model 
```{r}
# Beta
yvar <- "na_percent"
df_model <- soil %>% 
  dplyr::select(one_of(c(yvar,"fecha", "estacion", "geo_parcela_nombre")))
form <- reformulate("fecha * estacion + (1|estacion:geo_parcela_nombre)", 
                    response = yvar)


library(glmmTMB)
m <- glmmTMB::glmmTMB(formula = form,
                         family = beta_family (link = "logit"),
                         data = df_model)

tabla_anova <- car::Anova(m)
tabla_anova$variable <- yvar
tabla_anova$factor <- rownames(tabla_anova)
assign(paste0("anova_", yvar), tabla_anova)

  tabla_anova %>% dplyr::select(fvalue = `Chisq`, p=`Pr(>Chisq)`, variable, factor) %>%  
  assign(paste0("anova_", yvar), ., inherits = TRUE)
car::Anova(m)
```

### *Post-hoc*
```{r}
tabla_postHoc(m)
```


# pH agua 
### Model 
```{r}
# Gamma
yvar <- "p_h_agua_eez"
df_model <- soil %>% 
  dplyr::select(one_of(c(yvar,"fecha", "estacion", "geo_parcela_nombre")))
form <- reformulate("fecha * estacion + (1|estacion:geo_parcela_nombre)", 
                    response = yvar)

m <- lme4::glmer(form, data = df_model, family= Gamma)
afex::mixed(form, data=df_model)
tabla_anova <- afex::mixed(form, data=df_model)
tabla_anova$anova_table
tabla_anova <- as.data.frame(tabla_anova$anova_table)
tabla_anova$variable <- yvar
tabla_anova$factor <- rownames(tabla_anova)

tabla_anova %>% dplyr::select(fvalue = `F`, p=`Pr(>F)`, variable, factor) %>%  
  assign(paste0("anova_", yvar), ., inherits = TRUE)

```

### *Post-hoc*
```{r}
tabla_postHoc(m)
```



# pH KCl 
### Model 
```{r}
# Gamma
yvar <- "p_h_k_cl"
df_model <- soil %>% 
  dplyr::select(one_of(c(yvar,"fecha", "estacion", "geo_parcela_nombre")))
form <- reformulate("fecha * estacion + (1|estacion:geo_parcela_nombre)", 
                    response = yvar)

m <- lme4::glmer(form, data = df_model, family= Gamma)
afex::mixed(form, data=df_model)
tabla_anova <- afex::mixed(form, data=df_model)
tabla_anova$anova_table
tabla_anova <- as.data.frame(tabla_anova$anova_table)
tabla_anova$variable <- yvar
tabla_anova$factor <- rownames(tabla_anova)

tabla_anova %>% dplyr::select(fvalue = `F`, p=`Pr(>F)`, variable, factor) %>%  
  assign(paste0("anova_", yvar), ., inherits = TRUE)

```

### *Post-hoc*
```{r}
tabla_postHoc(m)
```


# NH4 
- Prepare data
- We have only data for Autumn fire 
- The approach will be the following: Apply non-parametric Wilcoxon test to compare pre and postFire


```{r}
soilN <- raw_soil %>% 
  dplyr::select(fecha, fecha, estacion, zona, geo_parcela_nombre, 
                n_nh4,n_no3) %>% 
  drop_na() %>% 
  mutate(fecha = as.factor(fecha), 
         estacion = as.factor(estacion)
  )
  
xtabs(~estacion+fecha, data = soilN)

```
### Model 
```{r}
# Wilcoxon
yvar <- "n_nh4"
df_model <- soilN %>% 
  dplyr::select(one_of(c(yvar, "fecha", "zona", "geo_parcela_nombre")))
form <- reformulate("fecha", response = yvar)
m <- wilcox.test(form, data=df_model)

tabla_anova <- data.frame(
  fvalue = m$statistic, 
  p = m$p.value, 
  variable = yvar,
  factor = "fecha"
)

tabla_anova %>%  
  assign(paste0("anova_", yvar), ., inherits = TRUE)

```


# NO3
### Model 
```{r}
# Wilcoxon
yvar <- "n_no3"
df_model <- soilN %>% 
  dplyr::select(one_of(c(yvar, "fecha", "zona", "geo_parcela_nombre")))
form <- reformulate("fecha", response = yvar)
m <- wilcox.test(form, data=df_model)

tabla_anova <- data.frame(
  fvalue = m$statistic, 
  p = m$p.value, 
  variable = yvar,
  factor = "fecha"
)

tabla_anova %>%  
  assign(paste0("anova_", yvar), ., inherits = TRUE)

```




# General Overview
## Mean + SE table 
```{r}
library(plotrix)
library(gtsummary)
soil %>% 
  dplyr::select(estacion, fecha, one_of("humedad","n_nh4","n_no3","fe_percent","k_percent","mg_percent","na_percent","n_percent","c_percent","c_n","cic","p","mo","p_h_k_cl","p_h_agua_eez")) %>% 
  tbl_strata(
    strata = fecha, 
    .tbl_fun = 
      ~.x %>% 
      tbl_summary(
        by = estacion,
        missing = "no",
        type = c(p) ~ "continuous",
       # label = list(ab ~ "Juvenile abundance"),
        statistic = all_continuous() ~ "{mean} ({std.error})",
        digits = everything() ~ 2
      ))
```

```{r}
library(plotrix)
library(gtsummary)
soil %>% 
  dplyr::select(estacion, fecha, one_of("humedad","n_nh4","n_no3","fe_percent","k_percent","mg_percent","na_percent","n_percent","c_percent","c_n","cic","p","mo","p_h_k_cl","p_h_agua_eez")) %>% 
  tbl_strata(
    strata = estacion, 
    .tbl_fun = 
      ~.x %>% 
      tbl_summary(
        by = fecha,
        missing = "no",
        type = c(p) ~ "continuous",
       # label = list(ab ~ "Juvenile abundance"),
        statistic = all_continuous() ~ "{mean} ({std.error})",
        digits = everything() ~ 2
      ))
```

## Figures 
```{r}
pos <- position_dodge(0.15)

medias <- soil %>% dplyr::select(
  fecha, estacion, humedad:p_h_agua_eez)%>% 
  group_by(fecha, estacion) %>% 
  summarise_at(vars(humedad:p_h_agua_eez), ~mean(., na.rm=TRUE)) %>% 
  pivot_longer(humedad:p_h_agua_eez, names_to = "var", values_to = "mean")

errores <- soil %>% dplyr::select(
  fecha, estacion, humedad:p_h_agua_eez)%>% 
  group_by(fecha, estacion) %>% 
  summarise_at(vars(humedad:p_h_agua_eez), ~plotrix::std.error(., na.rm=TRUE)) %>% 
  pivot_longer(humedad:p_h_agua_eez, names_to = "var", values_to = "se")

medias <- medias %>% dplyr::inner_join(errores) %>% 
  mutate(varname = case_when(
    var == "n_percent" ~ "N",
    var == "p" ~ "P",
    var == "k_percent" ~ "K",
    var == "mo" ~ "M.O.",
    var == "c_percent" ~ "C",
    var == "c_n" ~ "C/N",
    var == "humedad" ~ "Humedad",
    var == "n_nh4" ~ "NH4",
    var == "fe_percent" ~ "Fe",
    var == "mg_percent" ~ "Mg",
    var == "na_percent" ~ "Na",
    var == "cic" ~ "C.I.C.",
    var == "n_no3" ~ "NO3",
    var == "p_h_k_cl"  ~ "pH (KCl)",
    var == "p_h_agua_eez" ~ "pH (Agua)",
    var == "p_h_agua" ~ "pH (Agua)_old"
  ))

plotsuelos <- function(df, ncol, nrow){
    ggplot(df, aes(x=fecha, y = mean, 
               color = estacion, shape = estacion, group=estacion)) + 
    geom_point(position = pos, size = 2) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se), 
                  position = pos, width = .01) +
    geom_line(position = pos) +
    theme_bw() +
    theme(strip.background = element_rect(fill="white")) + 
    ylab("") + xlab("") +
  facet_wrap(~varname, scales = "free_y", ncol=ncol)
}
```


```{r}
medias %>% 
  filter(var %in% c("n_percent", "p", "k_percent",
                    "mo", "c_percent", "c_n")) %>% 
  plotsuelos(ncol=2) 
```

```{r}
medias %>% 
  filter(var %in% c("cic", "mg_percent", 
                    "na_percent", "humedad")) %>% 
  plotsuelos(ncol=2)
```

```{r}
medias %>% 
  filter(var %in% c("n_nh4","n_no3",
                    "p_h_k_cl", "p_h_agua_eez")) %>% 
  plotsuelos(ncol=2) 
```

## Anovas table
```{r}

anovas <- bind_rows(anova_c_n, 
                    anova_cic,
                    anova_k_percent,
                    anova_mg_percent,
                    anova_mo,
                    anova_n_nh4,
                    anova_n_no3, 
                    anova_p, 
                    anova_p_h_agua_eez, 
                    anova_p_h_k_cl, 
                    anova_humedad, 
                    anova_n_percent, 
                    anova_c_percent, 
                    anova_na_percent) %>% 
  relocate(variable, factor)
rownames(anovas) <- NULL
```

```{r}
library(kableExtra)

anovas %>% 
  mutate(across(where(is.numeric), round, 3)) %>% 
  pivot_wider(names_from = factor, values_from = c(fvalue, p)) %>% 
  relocate(variable,
           fvalue_estacion, p_estacion, 
           fvalue_fecha, p_fecha, 
           `fvalue_fecha:estacion`, `p_fecha:estacion`
           ) %>% 
  kbl(col.names = 
        c("Variables", rep(c("F", "p"),3))) %>% 
  add_header_above(c("", "estacion" = 2, "fecha" = 2, "estacion x fecha" = 2)) %>% 
  kable_styling() 

```

```{r}
anovas %>% 
  mutate(across(where(is.numeric), round, 3)) %>% 
  pivot_wider(names_from = factor, values_from = c(fvalue, p)) %>% 
  relocate(variable,
           fvalue_estacion, p_estacion, 
           fvalue_fecha, p_fecha, 
           `fvalue_fecha:estacion`, `p_fecha:estacion`
           ) %>% 
  write_csv(here::here("output/anovas_pre_post_epoca.csv"))
```

# Gráficos feos feísimos
```{r}
mediasFeas <- medias %>% 
  unite("fac", estacion:fecha, remove = FALSE)

plotfeos <- function(df){
  coloritos_feos <- c("darkgreen","darkred")
  
  ggplot(df, aes(x=fecha, y = mean, 
               color = fecha, fill= fecha, group=fecha)) + 
    geom_bar(stat="identity", width = .5) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se), 
                  position = pos, width = .3) +
    theme_bw() + 
  scale_fill_manual(values=coloritos_feos) + 
  scale_color_manual(values=coloritos_feos) +
  facet_grid(var~estacion, 
             switch = "y", scales = "free_y") + 
  theme(strip.background = element_rect(fill="white"), 
        legend.position = "none",
        strip.placement = "outside") + 
    ylab("") + xlab("")}

```

```{r}
mediasFeas %>% 
  filter(var %in% c("n_percent", "p", "k_percent")) %>% 
  plotfeos()
```



```{r}
mediasFeas %>% 
  filter(var %in% c("mo", "c_percent", "c_n")) %>% 
  plotfeos()
```


```{r}
mediasFeas %>% 
  filter(var %in% c("cic", "mg_percent", 
                    "na_percent")) %>% 
 plotfeos()
```

```{r}
mediasFeas %>% 
  filter(var %in% c(
                    "p_h_k_cl", "p_h_agua_eez", "humedad")) %>% 
  plotfeos()
```


```{r}
mediasFeas %>% 
  filter(var %in% c("n_nh4","n_no3")) %>% 
  plotfeos()
```

# Boostrapping methods

- Get mean and IC 95% by boostraping 

```{r}
b <- soil %>% 
  dplyr::select(estacion, fecha, one_of("humedad","n_nh4","n_no3","fe_percent","k_percent","mg_percent","na_percent","n_percent","c_percent","c_n","cic","p","mo","p_h_k_cl","p_h_agua_eez")) 
```

- We generate a function to compute the mean and the ci (both by bootstrapping) for each  soil variable and groupped by estacion and fecha. 
- Then we generated a dataframe with this info and plot 

```{r}
b <- soil %>% 
  dplyr::select(
    estacion, fecha,
    one_of("humedad","n_nh4","n_no3","fe_percent","k_percent",
           "mg_percent","na_percent","n_percent",
           "c_percent","c_n","cic","p","mo","p_h_k_cl","p_h_agua_eez")) %>%
  group_by(estacion, fecha) %>% tidyr::nest()

# visualiza los datos
# lapply(b$data, head)

## Generate a custom function to compute bootstrap 
bootme <- function(df, myvar){
  require(magrittr)
  require(purrr)
  require(boot)
  require(tidyverse)
  
  ## boot_mean
boot_mean <- function(d, i) {
    mean(d[i], na.rm=TRUE)
}


  df %<>%
  dplyr::mutate(
    booted =
      purrr::map(
        .x = data, # The list-column containing <S3: tibble>
        ~ boot::boot(
          data = .x[[myvar]], # The <S3 tibble> column being sampled
          statistic = boot_mean, # The user-defined function
          R = 1000, # The number of replicates
          stype = "i"
        )
      ),
    booted_ci = purrr::map(
      .x = booted, 
      ~ boot::boot.ci(.x,
                      conf = 0.95, # Interval width
                      type = "bca"))  # Calculate a BCa interval
    ) 

# tidy
out <- df %>%
    dplyr::mutate(
      variable = myvar, 
      statistic = purrr::map(.x = booted_ci, ~ .x$t0), # mean point estimates
      lower_ci = purrr::map(.x = booted_ci, ~ .x$bca[[4]]), # lower 2.5% limit
      upper_ci = purrr::map(.x = booted_ci, ~ .x$bca[[5]])) %>% # upper 97.5% limit
    dplyr::select(-data, -booted, -booted_ci) %>%
    tidyr::unnest(cols = c(statistic, lower_ci, upper_ci))

return(out)
} 

vi <- c("humedad","fe_percent","k_percent","mg_percent","na_percent","n_percent","c_percent","c_n","cic","p","mo","p_h_k_cl","p_h_agua_eez")

# "n_nh4","n_no3"

bm <- c()
for(i in 1:length(vi)){
  aux <- bootme(b, vi[i])
  bm <- rbind(bm, aux)
} 

# For NO3 and Nh4 select 
bn <- soil %>% 
  dplyr::select(fecha, n_nh4, n_no3) %>%
  group_by(fecha) %>% tidyr::nest()


bmm <- bind_rows(
  bm, 
  bootme(bn, "n_nh4") %>% mutate(estacion = "Ot"),
  bootme(bn, "n_no3") %>% mutate(estacion = "Ot")
) 

write_csv(bmm, here::here("output/meanboot_pre_post_epoca.csv"))

pos <- position_dodge(0.15)


plotCI <- function(df, ncol, nrow) {
  ggplot(df, aes(
    x = fecha, y = statistic,
    color = estacion, shape = estacion, group = estacion
  )) +
    geom_linerange(aes(ymin = lower_ci, ymax = upper_ci), position = pos) +
    geom_line(position = pos) +
    geom_point(position = pos, size = 2) +
    theme_bw() +
    theme(strip.background = element_rect(fill = "white")) +
    ylab("") +
    xlab("") +
    facet_wrap(~variable, scales = "free_y")
}
```


## Plots with CI 95 % (bootstraping bca)

```{r}
bmm %>% 
  filter(variable %in% c("n_percent", "p", "k_percent",
                    "mo", "c_percent", "c_n")) %>% 
  plotCI()
```

```{r}
bmm %>% 
  filter(variable %in% c("cic", "mg_percent", 
                    "na_percent", "humedad")) %>% 
  plotCI()
```


```{r}
bmm %>% 
  filter(variable %in% c("n_nh4","n_no3",
                    "p_h_k_cl", "p_h_agua_eez")) %>% 
  plotCI()
```


