---
title: "Resilience of Soil propierties"
author: "ajpelu"
date: "2021-09-08"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(knitr)
library(rmdformats)

## Global options
options(max.print="100")
opts_chunk$set(echo=FALSE,
	             cache=FALSE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE,
               autodep = TRUE)
```
# Introduction
- Analysis of resilience of soil propierties. 
- Only for Autumn treatment (*i.e.* zona == "P"; zona == "NP") 
- Interpret zona as "grazing effect": 
  - zona == "P" corresponds to *Browsing*
  - zona == "NP" corresponds to *No Browsing*


```{r pkgs}
library("tidyverse")
library("here")
library("janitor")
library("ggpubr")
library("fitdistrplus")
library("lme4")
library("lmerTest")
library("emmeans")
library("multcomp")
library("performance")
library("afex")
library("DHARMa")
library("glmmTMB")
library("glmmADMB")
library("car")
```

## Prepare data 
```{r prepare-data, echo=TRUE}
raw_soil <- readxl::read_excel(
  here::here("data/Resultados_Suelos_2018_2021_v2.xlsx"),
  sheet = "SEGUIMIENTO_SUELOS_sin_ouliers"
) %>% 
  janitor::clean_names() %>% 
  mutate(
    treatment_name =
      case_when(
        str_detect(geo_parcela_nombre, "NP_") ~ "Autumn Burning / No Browsing",
        str_detect(geo_parcela_nombre, "PR_") ~ "Spring Burning / Browsing",
        str_detect(geo_parcela_nombre, "P_") ~ "Autumn Burning / Browsing"
      ),
    zona =
      case_when(
        str_detect(geo_parcela_nombre, "NP_") ~ "QOt_NP",
        str_detect(geo_parcela_nombre, "PR_") ~ "QPr_P",
        str_detect(geo_parcela_nombre, "P_") ~ "QOt_P"
      ),
    fecha = lubridate::ymd(fecha),
    pre_post_quema = case_when(
      pre_post_quema == "Prequema" ~ "0 preQuema",
      pre_post_quema == "Postquema" ~ "1 postQuema"
    )
  ) 

```

- Compute date as months after fire 

```{r, echo=TRUE}
autumn_fire <- lubridate::ymd("2018-12-18")

soil <- raw_soil %>% 
  filter(zona != "QPr_P") %>% 
  mutate(zona = as.factor(zona)) %>% 
  mutate(meses = as.factor(
    case_when(
    fecha == "2018-12-11"  ~ as.character("-1"), 
    fecha != "2018-12-11"  ~ as.character(lubridate::interval(autumn_fire, lubridate::ymd(fecha)) %/% months(1))
  ))) %>% 
  mutate(pastoreo = as.factor(
    case_when(
    zona == "QOt_P" ~ "Browsing",
    zona == "QOt_NP" ~ "No Browsing"
  ))) %>% 
  relocate(pastoreo, fecha, meses) %>% 
  dplyr::select(-pre_post_quema, -tratamiento)

xtabs(~meses+pastoreo, data = soil)
```

## Compute resilience 

For each soil variable we compute the ***Resilience*** value according to the following equation (*sensu* LLoret et al. 2011): 
$$Resilience = preFire / postFire$$
- We computed the resilience value for each time step: just after fire (month=0) and after 22 and 29 months. 

- The we groupped soil variables by treatment (Browsing *vs* No Browsing) 

- Finally we explored graphically the results. 

```{r, echo = TRUE}
computeResilience <- function(df, variable){ 
  df %>% 
  mutate(replica = str_remove(geo_suelos_nombre, "Q|_Q")) %>% 
  dplyr::select(one_of("pastoreo", "meses", "geo_parcela_nombre",
                "replica", variable)) %>% 
    pivot_wider(values_from = one_of(variable), 
                names_from = meses, names_prefix = "t") %>% 
    filter(!(str_detect(replica,"E3"))) %>% 
  rename(control = `t-1`) %>% 
  # resilience r = post/pre
  mutate(res.0 = t0/control,
         res.22 = t22/control, 
         res.29 = t29/control) %>% 
  dplyr::select(pastoreo, 
                geo_parcela_nombre, 
                res.0, res.22, res.29) %>% 
  pivot_longer(res.0:res.29, names_to = "resilience") %>% 
    mutate(myvar = variable)
  }


x <- bind_rows(computeResilience(soil, "humedad"),
computeResilience(soil, "n_nh4"),
computeResilience(soil, "n_no3"),
computeResilience(soil, "fe_percent"),
computeResilience(soil, "k_percent"),
computeResilience(soil, "mg_percent"),
computeResilience(soil, "na_percent"),
computeResilience(soil, "n_percent"),
computeResilience(soil, "c_percent"),
computeResilience(soil, "c_n"),
computeResilience(soil, "cic"),
computeResilience(soil, "p"),
computeResilience(soil, "mo"),
computeResilience(soil, "p_h_agua_eez"),
computeResilience(soil, "p_h_k_cl"))


x <- x %>% mutate(meses = str_remove(resilience, "res."))
```

# Plots

```{r}

plotResilience <- function(df){  
  ggplot(df, aes(x=meses, y=value, colour=pastoreo)) + 
  geom_boxplot() + 
  ylab("Resilience") + xlab("Months after fire") + 
  facet_wrap(~myvar, scales = "free_y") + 
  geom_hline(yintercept = 1) +
  scale_color_manual(values=c("blue", "darkgreen")) + 
  theme_bw()
} 
```


```{r}
x %>% filter(value < 6) %>% 
  filter(!(myvar %in% c("n_nh4", "n_no3"))) %>% 
  filter(myvar %in% c("n_percent", "p", "k_percent",
                    "mo", "c_percent", "c_n")) %>% 
  plotResilience()
  
```


```{r}
x %>% filter(value < 6) %>% 
  filter(!(myvar %in% c("n_nh4", "n_no3"))) %>% 
  filter(myvar %in% c("cic", "mg_percent", 
                    "na_percent", "humedad",
                    "p_h_k_cl", "p_h_agua_eez")) %>% 
  plotResilience()
```




