

### Normality & Homocedasticity
```{r}








chequeo

chequeoDist(soil, "humedad")

%>% plot()

yvar <- "humedad"


df_model <- soil %>% 
  dplyr::select(one_of(c(yvar,"pre_post_quema", "zona", "geo_parcela_nombre")))
form <- reformulate("pre_post_quema * zona + (1|zona:geo_parcela_nombre)", 
                    response = yvar)
model <- lme4::lmer(form, data = df_model)

p.dist <- performance::check_distribution(model)



plot(performance::check_distribution(model)) 

print("Variances homogeneity?")
plot(performance::check_homogeneity(model))

print("Normality?")
performance::check_normality(model)

print("Normality, Random effects")
performance::check_normality(model, effects = "random")
```

### Model 
```{r}
m <- lme4::lmer(form, data = df_model)
anova(as_lmerModLmerTest(m))

tabla_anova <- as.data.frame(anova(as_lmerModLmerTest(m)))
tabla_anova$variable <- yvar
tabla_anova$factor <- rownames(tabla_anova)
tabla_anova

tabla_anova %>% dplyr::select(fvalue = `F value`, p=`Pr(>F)`, variable, factor) %>%  
  assign(paste0("anova_", yvar), ., inherits = TRUE)

```

```{r}
library(DHARMa)
simulationOutput <- simulateResiduals(fittedModel = m, plot = F)
plot(simulationOutput)
```


```{r}
# Ojo para lo de la anova en lmer ver esto:  https://stat.ethz.ch/pipermail/r-sig-mixed-models/2018q1/026596.html
```


### *Post-hoc*

```{r}
emmeans(m, list(pairwise ~ pre_post_quema), adjust = "tukey")
emmeans(m, list(pairwise ~zona), adjust = "tukey")
emmeans(m, list(pairwise ~ pre_post_quema | zona), adjust = "tukey")
```

```{r}
plot_interaction(df=soil, yvar="humedad")
```


# CIC 
### Check distribution 
```{r}
# Cullen & Frey plot 
descdist(soil$cic, boot = 1000)
```

### Normality & Homocedasticity
```{r}
yvar <- "cic"

df_model <- soil %>% 
  dplyr::select(one_of(c(yvar,"pre_post_quema", "zona", "geo_parcela_nombre")))
form <- reformulate("pre_post_quema * zona + (1|zona:geo_parcela_nombre)", 
                    response = yvar)
model <- lme4::lmer(form, data = df_model)

plot(performance::check_distribution(model)) 

print("Variances homogeneity?")
performance::check_homogeneity(model)

print("Normality?")
performance::check_normality(model)

print("Normality, Random effects")
performance::check_normality(model, effects = "random")
```

### Model 
```{r}
m <- lme4::lmer(form, data = df_model)
tabla_anova <- anova(as_lmerModLmerTest(m)) 
tabla_anova <- as.data.frame(tabla_anova)
tabla_anova$variable <- yvar
tabla_anova$factor <- rownames(tabla_anova)
tabla_anova %>% dplyr::select(fvalue = `F value`, p=`Pr(>F)`, variable, factor) %>%  
  assign(paste0("anova_", yvar), ., inherits = TRUE)
```

```{r}
# Ojo para lo de la anova en lmer ver esto:  https://stat.ethz.ch/pipermail/r-sig-mixed-models/2018q1/026596.html
```

### *Post-hoc*

```{r}
emmeans(m, list(pairwise ~ pre_post_quema), adjust = "tukey")
emmeans(m, list(pairwise ~zona), adjust = "tukey")
emmeans(m, list(pairwise ~ pre_post_quema | zona), adjust = "tukey")
```

```{r}
plot_interaction(df=soil, yvar="cic")
```


# % C 
### Check distribution 
```{r}
# Cullen & Frey plot 
descdist(soil$c_percent, boot = 1000)
```

### Normality & Homocedasticity
```{r}
yvar <- "c_percent"

df_model <- soil %>% 
  dplyr::select(one_of(c(yvar,"pre_post_quema", "zona", "geo_parcela_nombre")))
form <- reformulate("pre_post_quema * zona + (1|zona:geo_parcela_nombre)", 
                    response = yvar)
model <- lme4::lmer(form, data = df_model)

plot(performance::check_distribution(model)) 

print("Variances homogeneity?")
performance::check_homogeneity(model)

print("Normality?")
performance::check_normality(model)

print("Normality, Random effects")
performance::check_normality(model, effects = "random")
```

### Model 
```{r}
m <- lme4::lmer(form, data = df_model)
anova(as_lmerModLmerTest(m))

tabla_anova <- as.data.frame(anova(as_lmerModLmerTest(m)))
tabla_anova$variable <- yvar
tabla_anova$factor <- rownames(tabla_anova)

tabla_anova %>% dplyr::select(fvalue = `F value`, p=`Pr(>F)`, variable, factor) %>%  
  assign(paste0("anova_", yvar), ., inherits = TRUE)

```


### *Post-hoc*

```{r}
emmeans(m, list(pairwise ~ pre_post_quema), adjust = "tukey")
emmeans(m, list(pairwise ~zona), adjust = "tukey")
emmeans(m, list(pairwise ~ pre_post_quema | zona), adjust = "tukey")
```

```{r}
plot_interaction(df=soil, yvar="c_percent")
```


# % Fe 
### Check distribution 
```{r}
# Cullen & Frey plot 
descdist(soil$fe_percent, boot = 1000)
```

### Normality & Homocedasticity
```{r}
yvar <- "fe_percent"

df_model <- soil %>% 
  dplyr::select(one_of(c(yvar,"pre_post_quema", "zona", "geo_parcela_nombre")))
form <- reformulate("pre_post_quema * zona + (1|zona:geo_parcela_nombre)", 
                    response = yvar)
model <- lme4::lmer(form, data = df_model)

plot(performance::check_distribution(model)) 

print("Variances homogeneity?")
performance::check_homogeneity(model)

print("Normality?")
performance::check_normality(model)

print("Normality, Random effects")
performance::check_normality(model, effects = "random")
```

### Model 
- Gamma 
```{r}
# See here https://zian999.github.io/posts/2019/lrt_pvalues_for_glmer/ 
m <- lme4::glmer(form, data = df_model, family= Gamma)
afex::mixed(form, data=df_model)
tabla_anova <- afex::mixed(form, data=df_model)
tabla_anova$anova_table
tabla_anova <- as.data.frame(tabla_anova$anova_table)
tabla_anova$variable <- yvar
tabla_anova$factor <- rownames(tabla_anova)

tabla_anova %>% dplyr::select(fvalue = `F`, p=`Pr(>F)`, variable, factor) %>%  
  assign(paste0("anova_", yvar), ., inherits = TRUE)

```

### *Post-hoc*

```{r}
emmeans(m, list(pairwise ~ pre_post_quema), adjust = "tukey")
emmeans(m, list(pairwise ~zona), adjust = "tukey")
emmeans(m, list(pairwise ~ pre_post_quema | zona), adjust = "tukey")
```

```{r}
plot_interaction(df=soil, yvar="fe_percent")
```


# % K 
### Check distribution 
```{r}
# Cullen & Frey plot 
descdist(soil$k_percent, boot = 1000)
```

### Normality & Homocedasticity
```{r}
yvar <- "k_percent"

df_model <- soil %>% 
  dplyr::select(one_of(c(yvar,"pre_post_quema", "zona", "geo_parcela_nombre")))
form <- reformulate("pre_post_quema * zona + (1|zona:geo_parcela_nombre)", 
                    response = yvar)
model <- lme4::lmer(form, data = df_model)

plot(performance::check_distribution(model)) 

print("Variances homogeneity?")
performance::check_homogeneity(model)

print("Normality?")
performance::check_normality(model)

print("Normality, Random effects")
performance::check_normality(model, effects = "random")
```

### Model 
- Gamma 
```{r}
# See here https://zian999.github.io/posts/2019/lrt_pvalues_for_glmer/ 
m <- lme4::glmer(form, data = df_model, family= Gamma)

tabla_anova <- afex::mixed(form, data=df_model)
tabla_anova$anova_table
tabla_anova <- as.data.frame(tabla_anova$anova_table)
tabla_anova$variable <- yvar
tabla_anova$factor <- rownames(tabla_anova)

  tabla_anova %>% dplyr::select(fvalue = `F`, p=`Pr(>F)`, variable, factor) %>%  
  assign(paste0("anova_", yvar), ., inherits = TRUE)


```

### *Post-hoc*

```{r}
emmeans(m, list(pairwise ~ pre_post_quema), adjust = "tukey")
emmeans(m, list(pairwise ~ zona), adjust = "tukey")
emmeans(m, list(pairwise ~ pre_post_quema | zona), adjust = "tukey")
```

```{r}
plot_interaction(df=soil, yvar="k_percent")
```


# % Mg 
### Check distribution 
```{r}
# Cullen & Frey plot 
descdist(soil$mg_percent, boot = 1000)
```

### Normality & Homocedasticity
```{r}
yvar <- "mg_percent"

df_model <- soil %>% 
  dplyr::select(one_of(c(yvar,"pre_post_quema", "zona", "geo_parcela_nombre")))
form <- reformulate("pre_post_quema * zona + (1|zona:geo_parcela_nombre)", 
                    response = yvar)
model <- lme4::lmer(form, data = df_model)

plot(performance::check_distribution(model)) 

print("Variances homogeneity?")
performance::check_homogeneity(model)

print("Normality?")
performance::check_normality(model)

print("Normality, Random effects")
performance::check_normality(model, effects = "random")
```

### Model 
- Gamma 
```{r}
# See here https://zian999.github.io/posts/2019/lrt_pvalues_for_glmer/ 
m <- lme4::glmer(form, data = df_model, family= Gamma)
tabla_anova <- afex::mixed(form, data=df_model)
tabla_anova$anova_table
tabla_anova <- as.data.frame(tabla_anova$anova_table)
tabla_anova$variable <- yvar
tabla_anova$factor <- rownames(tabla_anova)

  tabla_anova %>% dplyr::select(fvalue = `F`, p=`Pr(>F)`, variable, factor) %>%  
  assign(paste0("anova_", yvar), ., inherits = TRUE)


```

### *Post-hoc*

```{r}
emmeans(m, list(pairwise ~ pre_post_quema), adjust = "tukey")
emmeans(m, list(pairwise ~ zona), adjust = "tukey")
emmeans(m, list(pairwise ~ pre_post_quema | zona), adjust = "tukey")
```

```{r}
plot_interaction(df=soil, yvar="mg_percent")
```

# C/N 
### Check distribution 
```{r}
# Cullen & Frey plot 
descdist(soil$c_n, boot = 1000)
```

### Normality & Homocedasticity
```{r}
yvar <- "c_n"

df_model <- soil %>% 
  dplyr::select(one_of(c(yvar,"pre_post_quema", "zona", "geo_parcela_nombre")))
form <- reformulate("pre_post_quema * zona + (1|zona:geo_parcela_nombre)", 
                    response = yvar)
model <- lme4::lmer(form, data = df_model)

plot(performance::check_distribution(model)) 

print("Variances homogeneity?")
performance::check_homogeneity(model)

print("Normality?")
performance::check_normality(model)

print("Normality, Random effects")
performance::check_normality(model, effects = "random")
```

### Model 
- Gamma 
```{r}
# See here https://zian999.github.io/posts/2019/lrt_pvalues_for_glmer/ 
m <- lme4::glmer(form, data = df_model, family= Gamma)
tabla_anova <- afex::mixed(form, data=df_model)
tabla_anova$anova_table
tabla_anova <- as.data.frame(tabla_anova$anova_table)
tabla_anova$variable <- yvar
tabla_anova$factor <- rownames(tabla_anova)

  tabla_anova %>% dplyr::select(fvalue = `F`, p=`Pr(>F)`, variable, factor) %>%  
  assign(paste0("anova_", yvar), ., inherits = TRUE)



```

### *Post-hoc*

```{r}
emmeans(m, list(pairwise ~ pre_post_quema), adjust = "tukey")
emmeans(m, list(pairwise ~ zona), adjust = "tukey")
emmeans(m, list(pairwise ~ pre_post_quema | zona), adjust = "tukey")
```

```{r}
plot_interaction(df=soil, yvar="c_n")
```

# P 
### Check distribution 
```{r}
# Cullen & Frey plot 
descdist(soil$p, boot = 1000)
```

### Normality & Homocedasticity
```{r}
yvar <- "p"

df_model <- soil %>% 
  dplyr::select(one_of(c(yvar,"pre_post_quema", "zona", "geo_parcela_nombre")))
form <- reformulate("pre_post_quema * zona + (1|zona:geo_parcela_nombre)", 
                    response = yvar)
model <- lme4::lmer(form, data = df_model)

plot(performance::check_distribution(model)) 

print("Variances homogeneity?")
performance::check_homogeneity(model)

print("Normality?")
performance::check_normality(model)

print("Normality, Random effects")
performance::check_normality(model, effects = "random")
```

### Model 
- Negative Binomial 
```{r}
# See here https://zian999.github.io/posts/2019/lrt_pvalues_for_glmer/ 
m <- lme4::glmer.nb(form, data = df_model)
tabla_anova <- afex::mixed(form, data=df_model)
tabla_anova$anova_table
tabla_anova <- as.data.frame(tabla_anova$anova_table)
tabla_anova$variable <- yvar
tabla_anova$factor <- rownames(tabla_anova)

  tabla_anova %>% dplyr::select(fvalue = `F`, p=`Pr(>F)`, variable, factor) %>%  
  assign(paste0("anova_", yvar), ., inherits = TRUE)

```

### *Post-hoc*

```{r}
emmeans(m, list(pairwise ~ pre_post_quema), adjust = "tukey")
emmeans(m, list(pairwise ~ zona), adjust = "tukey")
emmeans(m, list(pairwise ~ pre_post_quema | zona), adjust = "tukey")
```

```{r}
plot_interaction(df=soil, yvar="p")
```


# MO 
### Check distribution 
```{r}
# Cullen & Frey plot 
descdist(soil$mo, boot = 1000)
```

### Normality & Homocedasticity
```{r}
yvar <- "mo"

df_model <- soil %>% 
  dplyr::select(one_of(c(yvar,"pre_post_quema", "zona", "geo_parcela_nombre")))
form <- reformulate("pre_post_quema * zona + (1|zona:geo_parcela_nombre)", 
                    response = yvar)
model <- lme4::lmer(form, data = df_model)

plot(performance::check_distribution(model)) 

print("Variances homogeneity?")
performance::check_homogeneity(model)

print("Normality?")
performance::check_normality(model)

print("Normality, Random effects")
performance::check_normality(model, effects = "random")
```

### Model 
- Gamma 
```{r}
# See here https://zian999.github.io/posts/2019/lrt_pvalues_for_glmer/ 
m <- lme4::glmer(form, data = df_model, family= Gamma)
tabla_anova <- afex::mixed(form, data=df_model)
tabla_anova$anova_table
tabla_anova <- as.data.frame(tabla_anova$anova_table)
tabla_anova$variable <- yvar
tabla_anova$factor <- rownames(tabla_anova)

  tabla_anova %>% dplyr::select(fvalue = `F`, p=`Pr(>F)`, variable, factor) %>%  
  assign(paste0("anova_", yvar), ., inherits = TRUE)


```

### *Post-hoc*

```{r}
emmeans(m, list(pairwise ~ pre_post_quema), adjust = "tukey")
emmeans(m, list(pairwise ~ zona), adjust = "tukey")
emmeans(m, list(pairwise ~ pre_post_quema | zona), adjust = "tukey")
```

```{r}
plot_interaction(df=soil, yvar="mo")
```





# % N 
### Check distribution 
```{r}
# Cullen & Frey plot 
descdist(soil$n_percent, boot = 1000)
```

### Normality & Homocedasticity
```{r}
yvar <- "n_percent"

df_model <- soil %>% 
  dplyr::select(one_of(c(yvar,"pre_post_quema", "zona", "geo_parcela_nombre")))
form <- reformulate("pre_post_quema * zona + (1|zona:geo_parcela_nombre)", 
                    response = yvar)
model <- lme4::lmer(form, data = df_model)

plot(performance::check_distribution(model)) 

print("Variances homogeneity?")
performance::check_homogeneity(model)

print("Normality?")
performance::check_normality(model)

#print("Normality, Random effects")
#performance::check_normality(model, effects = "random")
```

### Model 
- Beta (glmmADBM)
```{r, eval=FALSE}
# See here https://zian999.github.io/posts/2019/lrt_pvalues_for_glmer/ 
m <- lme4::glmer(form, data = df_model, family= Gamma)

df_model$geo_parcela_nombre <- as.factor(df_model$geo_parcela_nombre)

m <- glmmADMB::glmmadmb(
  n_percent ~ pre_post_quema * zona, 
  random= ~ 1 | zona/geo_parcela_nombre,
  data = df_model, family = "beta")
#tabla_anova <- car::Anova(m)
#assign(paste0("anova_", yvar), tabla_anova)
```

- Beta glmmTMB
```{r}
library(glmmTMB)
mtbm <- glmmTMB::glmmTMB(formula = form,
                         family = beta_family (link = "logit"),
                         data = df_model)

tabla_anova <- car::Anova(mtbm)
tabla_anova$variable <- yvar
tabla_anova$factor <- rownames(tabla_anova)
assign(paste0("anova_", yvar), tabla_anova)

  tabla_anova %>% dplyr::select(fvalue = `Chisq`, p=`Pr(>Chisq)`, variable, factor) %>%  
  assign(paste0("anova_", yvar), ., inherits = TRUE)


```


### *Post-hoc*

```{r}
# ojo 
m <- mtbm
emmeans(m, list(pairwise ~ pre_post_quema), adjust = "tukey")
emmeans(m, list(pairwise ~ zona), adjust = "tukey")
emmeans(m, list(pairwise ~ pre_post_quema | zona), adjust = "tukey")
```

```{r}
plot_interaction(df=soil, yvar="n_percent")
```


# % Na 
### Check distribution 
```{r}
# Cullen & Frey plot 
descdist(soil$na_percent, boot = 1000)
```

### Normality & Homocedasticity
```{r}
yvar <- "na_percent"

df_model <- soil %>% 
  dplyr::select(one_of(c(yvar,"pre_post_quema", "zona", "geo_parcela_nombre")))
form <- reformulate("pre_post_quema * zona + (1|zona:geo_parcela_nombre)", 
                    response = yvar)
model <- lme4::lmer(form, data = df_model)

plot(performance::check_distribution(model)) 

print("Variances homogeneity?")
performance::check_homogeneity(model)

print("Normality?")
performance::check_normality(model)

print("Normality, Random effects")
performance::check_normality(model, effects = "random")
```

### Model 
- Beta (glmmADBM)
```{r}
# See here https://zian999.github.io/posts/2019/lrt_pvalues_for_glmer/ 
df_model$geo_parcela_nombre <- as.factor(df_model$geo_parcela_nombre)

m <- glmmADMB::glmmadmb(
  na_percent ~ pre_post_quema * zona, 
  random= ~ 1 | zona/geo_parcela_nombre,
  data = df_model, family = "beta")
car::Anova(m)


ma <- afex::mixed(form, data=df_model)
afex::nice(ma)
```

- Beta glmmTMB
```{r}
library(glmmTMB)
mtbm <- glmmTMB::glmmTMB(formula = form,
                         family = beta_family (link = "logit"),
                         data = df_model)

tabla_anova <- as.data.frame(car::Anova(mtbm))
tabla_anova$variable <- yvar
tabla_anova$factor <- rownames(tabla_anova)

tabla_anova %>% dplyr::select(fvalue = `Chisq`, p=`Pr(>Chisq)`, variable, factor) %>%  
  assign(paste0("anova_", yvar), ., inherits = TRUE)
```


### *Post-hoc*

```{r}
# ojo 
# m <- mtbm
emmeans(m, list(pairwise ~ pre_post_quema), adjust = "tukey")
emmeans(m, list(pairwise ~ zona), adjust = "tukey")
emmeans(m, list(pairwise ~ pre_post_quema | zona), adjust = "tukey")
```

```{r}
plot_interaction(df=soil, yvar="na_percent")
```








# % p_h_agua_eez
### Check distribution 
```{r}
# Cullen & Frey plot 
descdist(soil$p_h_agua_eez, boot = 1000)
```

### Normality & Homocedasticity
```{r}
yvar <- "p_h_agua_eez"

df_model <- soil %>% 
  dplyr::select(one_of(c(yvar,"pre_post_quema", "zona", "geo_parcela_nombre")))
form <- reformulate("pre_post_quema * zona + (1|zona:geo_parcela_nombre)", 
                    response = yvar)
model <- lme4::lmer(form, data = df_model)

plot(performance::check_distribution(model)) 

print("Variances homogeneity?")
performance::check_homogeneity(model)

print("Normality?")
performance::check_normality(model)

print("Normality, Random effects")
performance::check_normality(model, effects = "random")
```

### Model 
- Gamma
```{r}
# See here https://zian999.github.io/posts/2019/lrt_pvalues_for_glmer/ 
df_model$geo_parcela_nombre <- as.factor(df_model$geo_parcela_nombre)

m <- lme4::glmer(form, data = df_model, family= Gamma)

tabla_anova <- afex::mixed(form, data=df_model)
tabla_anova$anova_table
tabla_anova <- as.data.frame(tabla_anova$anova_table)
tabla_anova$variable <- yvar
tabla_anova$factor <- rownames(tabla_anova)

  tabla_anova %>% dplyr::select(fvalue = `F`, p=`Pr(>F)`, variable, factor) %>%  
  assign(paste0("anova_", yvar), ., inherits = TRUE)
```


### *Post-hoc*

```{r}
# ojo 
emmeans(m, list(pairwise ~ pre_post_quema), adjust = "tukey")
emmeans(m, list(pairwise ~ zona), adjust = "tukey")
emmeans(m, list(pairwise ~ pre_post_quema | zona), adjust = "tukey")
```

```{r}
plot_interaction(df=soil, yvar="p_h_agua_eez")
```









# % p_h_k_cl
### Check distribution 
```{r}
# Cullen & Frey plot 
descdist(soil$p_h_k_cl, boot = 1000)
```

### Normality & Homocedasticity
```{r}
yvar <- "p_h_k_cl"

df_model <- soil %>% 
  dplyr::select(one_of(c(yvar,"pre_post_quema", "zona", "geo_parcela_nombre")))
form <- reformulate("pre_post_quema * zona + (1|zona:geo_parcela_nombre)", 
                    response = yvar)
model <- lme4::lmer(form, data = df_model)

plot(performance::check_distribution(model)) 

print("Variances homogeneity?")
performance::check_homogeneity(model)

print("Normality?")
performance::check_normality(model)

print("Normality, Random effects")
performance::check_normality(model, effects = "random")
```

### Model 
- Gamma 
```{r}
# See here https://zian999.github.io/posts/2019/lrt_pvalues_for_glmer/ 
df_model$geo_parcela_nombre <- as.factor(df_model$geo_parcela_nombre)

m <- lme4::glmer(form, data = df_model, family= Gamma)

tabla_anova <- afex::mixed(form, data=df_model)
tabla_anova$anova_table
tabla_anova <- as.data.frame(tabla_anova$anova_table)
tabla_anova$variable <- yvar
tabla_anova$factor <- rownames(tabla_anova)

  tabla_anova %>% dplyr::select(fvalue = `F`, p=`Pr(>F)`, variable, factor) %>%  
  assign(paste0("anova_", yvar), ., inherits = TRUE)

```


### *Post-hoc*

```{r}
# ojo 
emmeans(m, list(pairwise ~ pre_post_quema), adjust = "tukey")
emmeans(m, list(pairwise ~ zona), adjust = "tukey")
emmeans(m, list(pairwise ~ pre_post_quema | zona), adjust = "tukey")
```

```{r}
plot_interaction(df=soil, yvar="p_h_k_cl")
```



# NH4
- prepara datos 

```{r}
soilN <- raw_soil %>% 
  dplyr::select(zona, fecha, pre_post_quema, geo_parcela_nombre, 
                n_nh4, n_no3) %>% 
  drop_na() %>% 
  mutate(zona = as.factor(zona), 
         pre_post_quema = as.factor(pre_post_quema)
  )
  
xtabs(~pre_post_quema+zona, data = soilN)


soilN %>% 
  group_by(zona, pre_post_quema) %>% 
  summarise(N.nh4 = length(n_nh4),
            N.no3 = length(n_no3))
```

### Check distribution 
```{r}
# Cullen & Frey plot 
descdist(soilN$n_nh4, boot = 1000)
```

### Normality & Homocedasticity
```{r}
yvar <- "n_nh4"

df_model <- soilN %>% 
  dplyr::select(one_of(c(yvar,"pre_post_quema", "zona", "geo_parcela_nombre")))
form <- reformulate("pre_post_quema * zona + (1|zona:geo_parcela_nombre)", 
                    response = yvar)
model <- lme4::lmer(form, data = df_model)

plot(performance::check_distribution(model)) 

print("Variances homogeneity?")
performance::check_homogeneity(model)

print("Normality?")
performance::check_normality(model)

#print("Normality, Random effects")
#performance::check_normality(model, effects = "random")
```

### Model 
- Gamma 
```{r}
# See here https://zian999.github.io/posts/2019/lrt_pvalues_for_glmer/ 
m <- lme4::glmer(form, data = df_model, family= Gamma)
tabla_anova <- afex::mixed(form, data=df_model)
tabla_anova$anova_table
tabla_anova <- as.data.frame(tabla_anova$anova_table)
tabla_anova$variable <- yvar
tabla_anova$factor <- rownames(tabla_anova)

  tabla_anova %>% dplyr::select(fvalue = `F`, p=`Pr(>F)`, variable, factor) %>%  
  assign(paste0("anova_", yvar), ., inherits = TRUE)
```

### *Post-hoc*

```{r}
emmeans(m, list(pairwise ~ pre_post_quema), adjust = "tukey")
emmeans(m, list(pairwise ~ zona), adjust = "tukey")
emmeans(m, list(pairwise ~ pre_post_quema | zona), adjust = "tukey")
```

```{r}
plot_interaction(df=soilN, yvar="n_nh4")
```


# NO3
### Check distribution 
```{r}
# Cullen & Frey plot 
descdist(soilN$n_no3, boot = 1000)
```

### Normality & Homocedasticity
```{r}
yvar <- "n_no3"

df_model <- soilN %>% 
  dplyr::select(one_of(c(yvar,"pre_post_quema", "zona", "geo_parcela_nombre")))
form <- reformulate("pre_post_quema * zona + (1|zona:geo_parcela_nombre)", 
                    response = yvar)
model <- lme4::lmer(form, data = df_model)

plot(performance::check_distribution(model)) 

print("Variances homogeneity?")
performance::check_homogeneity(model)

print("Normality?")
performance::check_normality(model)

#print("Normality, Random effects")
#performance::check_normality(model, effects = "random")
```

### Model 
- Gamma 
```{r}
# See here https://zian999.github.io/posts/2019/lrt_pvalues_for_glmer/ 
m <- lme4::glmer(form, data = df_model, family= Gamma)
tabla_anova <- afex::mixed(form, data=df_model)
tabla_anova$anova_table
tabla_anova <- as.data.frame(tabla_anova$anova_table)
tabla_anova$variable <- yvar
tabla_anova$factor <- rownames(tabla_anova)

  tabla_anova %>% dplyr::select(fvalue = `F`, p=`Pr(>F)`, variable, factor) %>%  
  assign(paste0("anova_", yvar), ., inherits = TRUE)
```

### *Post-hoc*

```{r}
emmeans(m, list(pairwise ~ pre_post_quema), adjust = "tukey")
emmeans(m, list(pairwise ~ zona), adjust = "tukey")
emmeans(m, list(pairwise ~ pre_post_quema | zona), adjust = "tukey")
```

```{r}
plot_interaction(df=soilN, yvar="n_no3")
```



# Resumen general 
```{r, eval=FALSE}
library(plotrix)
library(gtsummary)
soil %>% 
  dplyr::select(zona, pre_post_quema, humedad:p_h_k_cl) %>% 
  tbl_strata(
    strata = pre_post_quema, 
    .tbl_fun = 
      ~.x %>% 
      tbl_summary(
        by = zona,
        missing = "no",
       # label = list(ab ~ "Juvenile abundance"),
        statistic = all_continuous() ~ "{mean} ({std.error})",
        digits = everything() ~ 2
      ))

```


```{r}
pos <- position_dodge(0.15)

medias <- soil %>% dplyr::select(
  pre_post_quema, zona, humedad:p_h_agua_eez)%>% 
  group_by(zona, pre_post_quema) %>% 
  summarise_at(vars(humedad:p_h_agua_eez), ~mean(., na.rm=TRUE)) %>% 
  pivot_longer(humedad:p_h_agua_eez, names_to = "var", values_to = "mean")

errores <- soil %>% dplyr::select(
  pre_post_quema, zona, humedad:p_h_agua_eez) %>% 
  group_by(zona, pre_post_quema) %>% 
  summarise_at(vars(humedad:p_h_agua_eez), ~plotrix::std.error(., na.rm=TRUE)) %>% 
  pivot_longer(humedad:p_h_agua_eez, names_to = "var", values_to = "se")

medias <- medias %>% dplyr::inner_join(errores) %>% 
  mutate(varname = case_when(
    var == "n_percent" ~ "N",
    var == "p" ~ "P",
    var == "k_percent" ~ "K",
    var == "mo" ~ "M.O.",
    var == "c_percent" ~ "C",
    var == "c_n" ~ "C/N",
    var == "humedad" ~ "Humedad",
    var == "n_nh4" ~ "NH4",
    var == "fe_percent" ~ "Fe",
    var == "mg_percent" ~ "Mg",
    var == "na_percent" ~ "Na",
    var == "cic" ~ "C.I.C.",
    var == "n_no3" ~ "NO3",
    var == "p_h_k_cl"  ~ "pH (KCl)",
    var == "p_h_agua_eez" ~ "pH (Agua)",
    var == "p_h_agua" ~ "pH (Agua)_old"
  ))

plotsuelos <- function(df, ncol, nrow){
    ggplot(df, aes(x=pre_post_quema, y = mean, 
               color = zona, shape = zona, group=zona)) + 
    geom_point(position = pos, size = 2) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se), 
                  position = pos, width = .01) +
    geom_line(position = pos) +
    theme_bw() +
    theme(strip.background = element_rect(fill="white")) + 
    ylab("") + xlab("") +
  facet_wrap(~varname, scales = "free_y", ncol=ncol)
}
```

```{r}
medias %>% 
  filter(var %in% c("n_percent", "p", "k_percent",
                    "mo", "c_percent", "c_n")) %>% 
  plotsuelos(ncol=2) 
```

```{r}
medias %>% 
  filter(var %in% c("cic", "mg_percent", 
                    "na_percent", "humedad")) %>% 
  plotsuelos(ncol=2)
```

```{r}
medias %>% 
  filter(var %in% c("n_nh4","n_no3",
                    "p_h_k_cl", "p_h_agua_eez")) %>% 
  plotsuelos(ncol=2) 
```



```{r}
# Anovas

anovas <- bind_rows(anova_c_n, 
                    anova_cic,
                    anova_k_percent,
                    anova_mg_percent,
                    anova_mo,
                    anova_n_nh4,
                    anova_n_no3, 
                    anova_p, 
                    anova_p_h_agua_eez, 
                    anova_p_h_k_cl, 
                    anova_humedad, 
                    anova_n_percent, 
                    anova_c_percent, 
                    anova_na_percent) %>% 
  relocate(variable, factor)
rownames(anovas) <- NULL
```

```{r}
anovas %>% write_csv(here::here("output/anovas_pre_post.csv"))
```

```{r}
library(kableExtra)

anovas %>% 
  mutate(across(where(is.numeric), round, 3)) %>% 
  pivot_wider(names_from = factor, values_from = c(fvalue, p)) %>% 
  relocate(variable,
           fvalue_zona, p_zona, 
           fvalue_pre_post_quema, p_pre_post_quema, 
           `fvalue_pre_post_quema:zona`, `p_pre_post_quema:zona`
           ) %>% 
  kbl(col.names = 
        c("Variables", rep(c("F", "p"),3))) %>% 
  add_header_above(c("", "zona" = 2, "fecha" = 2, "zona X fecha" = 2)) %>% 
  kable_styling() 

```



# Explore Distribution of the models 
```{r custom-functions}
chequeoDist <- function(df, response){ 
  df_model <- df %>% 
    dplyr::select(one_of(response,"pre_post_quema", "zona", "geo_parcela_nombre"))
  
  form <- reformulate("pre_post_quema * zona + (1|zona:geo_parcela_nombre)", 
                    response = response)
  model <- lme4::lmer(form, data = df_model)

  plot(performance::check_distribution(model))
  
  print("Variances homogeneity?")
  performance::check_homogeneity(model)
  print("Normality?")
  performance::check_normality(model)
}

```



# Humedad 
### Check distribution 
```{r}
response <- "humedad"
# Cullen & Frey plot 
descdist(soil$humedad, boot = 1000)
title(sub=response)
chequeoDist(soil, "humedad")
```










